Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-04-25 lqg-strings.tex
added a paragraph justifying the use of conformal gauge below eqn(34)
added citation for every 2d metric is conformally flat theorem
minor wording change in paragraph below eqn(34)
minor wording changes and addition of bimetric analysis to main file

Rewriting eq(\ref{eqn:corrected-NG-wave-equation}) in terms of $\sigma$ derivatives and parameterizing $\sigma$ as (let $F^+ = 1 + g \Delta / 2 (- h)$)
\begin{equation} \label{eqn:sigma-parameterization}
	\sigma (q) = \frac{1}{T} \int _{0} ^{q} \diff E \frac{1}{(F^+)^2} = \frac{E (q)}{T (F^+)^2} ,
\end{equation}
we find the wave equation
\begin{equation} \label{eqn:simplified-modified-wave-eqn}
	(F^+)^2 \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial ^2 \Vec{X}}{\partial \sigma ^2} = 0 ,
\end{equation}
from which we can see that the correction factor modifies the speed of propagation of the wave in the string as
\begin{equation} \label{eqn:wave-speed-corrected}
	v = \frac{c}{F^+}.
\end{equation}
This is in fact an expected result for a effective theory emerging from a discretised background geometry, see\cite{Bodendorfer-LQG}.

@@

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} \eta _{\lambda \nu} + \\
	+ \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} \eta _{\lambda \nu} = 0 ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. In conformal gauge $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal factor to fix the second term as $- m^2 X^{\mu}$, with $- m^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

@@

Following the general procedure, given in \cite{Callan1985Strings}, we can couple the worldsheet to a bulk metric as follows:
\begin{equation}\label{eqn:polyakov-curved}
	S_{P} = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu G_{\mu\nu},
\end{equation}
where $G_{\mu\nu}$ is now the bulk or target space metric. Using this form we can now write down the worldsheet Polyakov bimetric action for two bulk metrics:
\begin{align}\label{eqn:polyakov-flat}
	S_{PB} = & -\frac{T}{2} \int \diff^2 x\sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu G_{\mu\nu} \nonumber \\
	         & -\frac{T'}{2} \int \diff^2 x \sqrt{-h} h^{ab} \partial_a X^\mu \partial_b X^\nu H_{\mu\nu}.
\end{align}

@@

Following the usual procedure such as the one found in \cite{Tong2009Lectures} with the bulk metric fixed, we find two copies of the string energy-momentum tensor being set to zero, which can be used to find an expression for the induced metrics:
\begin{equation} \label{eqn:variation-wrt-g}
	\frac{\delta S_{P B}}{\delta g^{c d}} \overset{!}{=} 0
\end{equation}
\begin{multline} \label{eqn:g-em-tensor}
	T^{(G)} _{c d} := G_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\nu} - \\
	- \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = 0
\end{multline}

\begin{equation} \label{eqn:g-expression}
	g_{c d} = 2 f^{(G)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f^{(G)}$ is given by
\begin{equation} \label{eqn:g's-f-function}
	\frac{1}{f^{(G)}} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu},
\end{equation}
and completely analogously for the $H_{\mu \nu}$ metric,
\begin{equation} \label{eqn:variation-wrt-h}
	\frac{\delta S_{P B}}{\delta h^{c d}} \overset{!}{=} 0
\end{equation}
\begin{multline} \label{eqn:h-em-tensor}
	T^{(H)} _{c d} := H_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\mu} - \\
	- \frac{1}{2} h_{c d} h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = 0
\end{multline}
\begin{equation} \label{eqn:h-expression}
	h_{c d} = 2 f^{(H)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} H_{\mu \nu} ,
\end{equation}
with
\begin{equation} \label{eqn:h's-f-function}
	\frac{1}{f^{(H)}} = h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} .
\end{equation}

As for the embbeding fields equation, we get a non-homogeneous wave equation for the coupled metric
\begin{equation} \label{eqn:variation-wrt-X}
	\frac{\delta S_{P B}}{\delta X^{\lambda}} \overset{!}{=} 0
\end{equation}

>>>>

Rewriting eq(\ref{eqn:corrected-NG-wave-equation}) in terms of $\sigma$ derivatives and parameterizing $\sigma$ as (let $F^+ = 1 + g \Delta / 2 (- h)$)
\begin{equation} \label{eqn:sigma-parameterization}
	\sigma (q) = \frac{1}{T} \int _{0} ^{q} \diff E \frac{1}{(F^+)^2} ,
\end{equation}
we find the wave equation
\begin{equation} \label{eqn:simplified-modified-wave-eqn}
	(F^+)^2 \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial ^2 \Vec{X}}{\partial \sigma ^2} = 0 ,
\end{equation}
from which we can see that the correction factor modifies the speed of propagation of the wave in the string as
\begin{equation} \label{eqn:wave-speed-corrected}
	v = \frac{c}{F^+}.
\end{equation}
This is in fact an expected result for a effective theory emerging from a discretised background geometry, see\cite{Bodendorfer-LQG}.

@@

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} \eta _{\lambda \nu} + \\
	+ \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} \eta _{\lambda \nu} = 0 ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. In conformal gauge $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}
The reason we can still use conformal gauge even tho our action no longer has Weyl symmetry is because of a theorem stating that {\it every 2d metric is conformally flat}(theorem 7.2 in\cite{Schottenloher}).

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal factor to fix the second term as $- \mu ^2 X^{\mu}$, with $- \mu ^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

@@

Following the general procedure, given in \cite{Callan1985Strings}, we can couple the worldsheet to a bulk metric as follows:
\begin{equation}\label{eqn:polyakov-curved}
	S_{P} = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu G_{\mu\nu},
\end{equation}
where $G_{\mu\nu}$ is now the bulk or target space metric. Using this form we can now write down the worldsheet Polyakov bimetric action for two bulk metrics:
\begin{multline} \label{eqn:bimetric-Polyakov}
	S_{BP} = -\frac{T}{2} \int \diff^2 x\sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu G_{\mu\nu} - \\
	-\frac{T'}{2} \int \diff^2 x \sqrt{-h} h^{ab} \partial_a X^\mu \partial_b X^\nu H_{\mu\nu} + \\
	+ \int \diff ^2 x \Phi (g , h) \Psi (X) \Theta (G , H) ,
\end{multline}
where the last term defines an interaction term $S_{int}$.

@@

Following the usual procedure such as the one found in \cite{Tong2009Lectures} with the bulk metric fixed, we find two copies of the string energy-momentum tensor being set to zero, which can be used to find an expression for the induced metrics:
\begin{equation} \label{eqn:variation-wrt-g}
	\frac{\delta S_{B P}}{\delta g^{c d}} = 0
\end{equation}
\begin{multline} \label{eqn:g-em-tensor}
	T^{(G)} _{c d} := G_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\nu} - \\
	- \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T \sqrt{- g}} \frac{\delta S_{int}}{\delta g^{c d}}
\end{multline}

\begin{equation} \label{eqn:g-expression}
	g_{c d} = 2 f^{(G)} \left( \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \frac{2}{T \sqrt{- g}} \frac{\delta S_{int}}{\delta g^{c d}} \right) ,
\end{equation}
where the function $f^{(G)}$ is given by
\begin{equation} \label{eqn:g's-f-function}
	\frac{1}{f^{(G)}} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu},
\end{equation}
and completely analogously for the $H_{\mu \nu}$ metric,
\begin{equation} \label{eqn:variation-wrt-h}
	\frac{\delta S_{B P}}{\delta h^{c d}} = 0
\end{equation}
\begin{multline} \label{eqn:h-em-tensor}
	T^{(H)} _{c d} := H_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\mu} - \\
	- \frac{1}{2} h_{c d} h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T' \sqrt{- h}} \frac{\delta S_{int}}{\delta h^{c d}}
\end{multline}
\begin{equation} \label{eqn:h-expression}
	h_{c d} = 2 f^{(H)} \left( \partial _{c} X^{\mu} \partial _{d} X^{\nu} H_{\mu \nu} - \frac{2}{T' \sqrt{- h}} \frac{\delta S_{int}}{\delta h^{c d}} \right) ,
\end{equation}
with
\begin{equation} \label{eqn:h's-f-function}
	\frac{1}{f^{(H)}} = h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} .
\end{equation}
If $S_{int} = 0$ and the auxiliary metrics $g_{a b}$ and $h_{a b}$ coincide with the induced metrics from $G_{\mu \nu}$ and $H_{\mu \nu}$, then the action\eqref{eqn:bimetric-Polyakov} reduces to
\begin{equation} \label{eqn:reduced-bimetric-action}
	S' _{B P} = - T \int \diff ^2 x \sqrt{- g} - T' \int \diff ^2 x \sqrt{- h} ,
\end{equation}
which, if $T' = - T k \Delta / 2$ and $h = g^{-1}$ becomes precisely the Nambu-Goto area corrected action\eqref{eqn:modified-NG-action}. The condition $h = g^{-1}$ can be achieved by
\begin{equation} \label{eqn:h=SgS}
	h_{a b} = S_{a c} g^{c d} S_{d b} ,
\end{equation}
where $S \in \mathrm{SL}(2 , \mathbb{C})$. The fact that we require $S = S^{\dagger}$ for $h$ to be real and symmetric restricts $S$ to be a boost. By requiring that the induced metric by $H_{\mu \nu}$ be equal to\eqref{eqn:h=SgS}, we get
\begin{equation} \label{eqn:H-from-SgS}
	\partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} = S_{a c} g^{c d} S_{d b}
\end{equation}
from which follows
\begin{equation}
	H_{\mu \nu} = \frac{1}{8 (- g)} (\Tr (S \varepsilon))^2 G_{\mu \nu} ,
\end{equation}
constraining $S$ to necessarily being not symmetric otherwise the trace $\Tr (S \varepsilon)$ vanishes and we would get $H _{\mu \nu} = 0$. In fact, to recover the modified Nambu-Goto action\eqref{eqn:modified-NG-action}, we require $(\Tr (S \varepsilon))^2 = 8$, thus $H_{\mu \nu} = G_{\mu \nu} / (- g)$.

As for the embbeding fields equation, we get a non-homogeneous wave equation for the coupled metric
\begin{equation} \label{eqn:variation-wrt-X}
	\frac{\delta S_{P B}}{\delta X^{\lambda}} = 0
\end{equation}
\begin{equation} \label{eqn:bimetric-X-eom}
	\partial _{a} \left( F^{a b} _{\mu \lambda} \partial _{b} X^{\mu} \right) = \frac{1}{2} \partial _{\lambda} F^{a b} _{\mu \nu} \partial _{a} X^{\mu} \partial _{b} X^{\nu} ,
\end{equation}
where the coupled double-metric $F^{a b} _{\mu \nu}$ is given by
\begin{equation} \label{eqn:coupled-double-metric}
	F^{a b} _{\mu \nu} = T \sqrt{- g} g^{a b} G_{\mu \nu} + T' \sqrt{- h} h^{a b} H_{\mu \nu} .
\end{equation}
If the conditions for $h = g^{-1}$ are satisfied and $G_{\mu \nu} = \eta _{\mu \nu}$, eq\eqref{eqn:bimetric-X-eom} reduces to
\begin{multline} \label{eqn:bimetric-X-simplified-1}
	\partial _{a} \big( \big( \sqrt{- g} g^{a b} - \\
	- \frac{k \Delta}{2 (- g)} \sqrt{- g}^{-1} (S^{-1})^{a c} g_{c d} (S^{-1})^{d b} \big) \partial _{b} X^{\mu} \big) = 0 ,
\end{multline}
which in conformal gauge $g_{a b} = \phi (x) \eta _{a b}$ and with
\begin{equation} \label{eqn:S-simplification-condition}
	(S^{-1})^{a c} \eta _{c d} (S^{-1})^{d b} = \eta ^{a b}
\end{equation}
simplifies further to
\begin{equation} \label{eqn:bimetric-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}
This is the exact form derived earlier in\eqref{eqn:X-eom-simplified}, thus leading to the same Klein-Gordon equation.
